A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-13 paper/paper.Rnw
Add updated heatmap process and code

codedir <- file.path(wd, "code")
modeldir <- file.path(wd, "model")

@@

<<single-exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
archetype <- list.files(file.path(wd, "images", "consistency", "single-exemplar"), full.names = T)
pred_prob_plot(archetype, loaded_model, sort = F)
@

@@

<<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
exemplar <- list.files(file.path(wd, "images", "consistency", "multi-exemplar"), full.names = T)
pred_prob_plot(exemplar, loaded_model)
@

@@

<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
pred_prob_plot(rev(chevron), loaded_model, sort = F)
@

@@

<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
pred_prob_plot(rev(circle_text), loaded_model, sort = F)
@

@@

@@

Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}. Using these graphics, we can identify which locations in an image contribute most to the model's predictions. In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.

\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet. How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?} \svp{I think it's worth highlighting the fact that we'll never get a model that's perfect because that curly s is pretty much a circle, and there will always be weird shapes that partially activate different categories.}

\svp{We may want to explore using the border of the square to show whether the model actually assigns that label - like the bullet plots}
<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
heatmaps <- list.files(file.path(wd, "images", "heatmaps"),
                       pattern = "^heatmap", full.names = T)
knitr::include_graphics(heatmaps[grepl("chevron-circle", heatmaps)])
@

@@

\section{Conclusions}
\svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles. Such a method takes the ``impossible" problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution. A set of geometric class characteristics was defined to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis. The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.

>>>>

codedir <- file.path(wd, "code")
modeldir <- file.path(wd, "model")
imgdir <- file.path(wd, "images")

@@

<<eers, include = F>>=
eers <- purrr::map_dbl(aucs$data$eer, function(x){x$thresholds}) %>%
  setNames(classes)
@

<<single-exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
single_exemplars <- list.files(file.path(wd, "images", "consistency",
                                         "single-exemplar"), full.names = T)
pred_prob_plot(single_exemplars, loaded_model, sort = F, eer = eers)
@

@@

<<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
multi_exemplars <- list.files(file.path(wd, "images", "consistency", "multi-exemplar"), full.names = T)
pred_prob_plot(multi_exemplars, loaded_model, eer = eers)
@

@@

<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
pred_prob_plot(rev(chevron), loaded_model, sort = F, eer = eers)
@

@@

<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
pred_prob_plot(rev(circle_text), loaded_model, sort = F, eer = eers)
@

@@

\paragraph{Contrast}

<<low-contrast-consistency, fig.cap = "Images with low contrast, before and after correction", fig.pos = "p!", out.width = "100%", cache = T>>=
lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast"), full.names = T)
pred_prob_plot(lowcontrast, loaded_model, sort = F)
@

@@

Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}. Using these graphics, we can identify which locations in an image contribute most to the model's predictions. In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image. \fix{Many of these heatmaps illustrate that multiple classes share certain features, which understandably confuses the model.}

%\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet. How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?} \svp{I think it's worth highlighting the fact that we'll never get a model that's perfect because that curly s is pretty much a circle, and there will always be weird shapes that partially activate different categories.}

\fix{Susan, please generate these heatmaps and move them to the "static" folder. My computer still has issues with the transparent layer of the heatmap, so it's probably easiest to do what we did for my CC and have you generate the images so they turn out correctly. Once I can see them all, I'll decide which three heatmaps best illustrate the point that shape categories have similar features that (understandably) confuse the model and I'll write about those specific heatmaps.}

<<setup-heatmap>>=
imgs <- list.files(file.path(wd, "images", "heatmaps", "raw"), full.names = T)
@

<<generate-heatmaps, echo = F>>=
# My computer has troubles with the transparency of the overlay...
# This code is functional, but highest quality will come from generating these
# images on Bigfoot and loading them statically.

h_imgdir <- file.path(imgdir, "heatmaps", "processed")
if(!dir.exists(h_imgdir)) {dir.create(h_imgdir)}

calc_heatmap(imgs[which(grepl("chevron-circle", imgs))], loaded_model) %>%
  prune_heatmap(c("chevron", "quad", "circle")) %>%
  create_composite(save_file = T, outdir = h_imgdir)

calc_heatmap(imgs[which(grepl("corks", imgs))], loaded_model) %>%
  prune_heatmap(c("star", "text", "circle")) %>%
  create_composite(save_file = T, outdir = h_imgdir)

calc_heatmap(imgs[which(grepl("text", imgs))], loaded_model) %>%
  prune_heatmap(c("text", "circle", "line")) %>%
  create_composite(save_file = T, outdir = h_imgdir)

calc_heatmap(imgs[which(grepl("triangle", imgs))], loaded_model) %>%
  prune_heatmap(c("triangle", "quad", "chevron")) %>%
  create_composite(save_file = T, outdir = h_imgdir)

calc_heatmap(imgs[which(grepl("star", imgs))], loaded_model) %>%
  prune_heatmap(c("star", "bowtie", "polygon")) %>%
  create_composite(save_file = T, outdir = h_imgdir)

calc_heatmap(imgs[which(grepl("polygon", imgs))], loaded_model) %>%
  prune_heatmap(c("star", "chevron", "polygon")) %>%
  create_composite(save_file = T, outdir = h_imgdir)

heatmaps <- list.files(imgdir, pattern = "^heatmap")
@

<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
heatmaps <- list.files(file.path(wd, "images", "heatmaps", "static"),
                       pattern = "^heatmap", full.names = T)
knitr::include_graphics(heatmaps[grepl("chevron-circle", heatmaps)])
@

@@

\section{Conclusions}
\svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
\mt{I'll do this once Results is stable, so I can make it match the rest of the story.}
The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles. Such a method takes the ``impossible" problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution. A set of geometric class characteristics was defined to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis. The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.
